Missing links and the topological robustness of food webs

### 2022-09-20 manuscript/ms_C3_v2.tex
update

\title{Missing links, forbidden links, and the topological robustness of
food webs}

@@

\usepackage{lineno}
\linenumbers
\usepackage {amsmath}
\setlength\parindent{24pt}
\usepackage{setspace}\doublespacing
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\begin{document}
\maketitle

\begin{abstract}
\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\tightlist
\item
  Undersampling can lead to missing trophic interactions in recorded
  food webs, with potential consequences for the perceived functioning
  and stability of the food web. Undersampling can be compensated for by
  using food web models such as allometric diet breadth model (ADBM) to
  predict missing links. Simultaneously, models might predict links
  which cannot occur, i.e., false positives.
\item
  Previous research shows that (i) food web robustness (the inverse of
  the number of secondary extinctions occuring due to primary
  extinctions) increases with connectance (the number of trophic links
  divided by the number of possible links), and (ii) that predicted food
  webs usually have greater connectance than observed ones. Thus we
  expect that predicted food webs are more robust than observed ones.
  This expectation has never, to our knowledge, been tested, nor has the
  effect size been quantified.
\item
  We fill this research gap by comparing the robustness of observed food
  webs to the robustness of food webs predicted by a model (the ADBM)
  that can account for missing links, though can also make false
  positives. We did this for 12 different food webs from a wide variety
  of ecosystems.
\item
  We found, as expected, that the predicted food webs were more robust
  than the observed food webs, and this can be attributed to the higher
  connectance of the predicted food webs. On average, for every X unit
  of increase in connectance, we found the food webs to be robust by YY
  units for random extinction scenario.\textbf{OP: Here something about
  the effect size.}
\item
  These results show that undersampling can lead to large underestimates
  of food web robustness that can be compensated for by filling in
  missing links with food web models. Nevertheless, increased
  connectance may contribute to lower dynamical stability, and so it
  would be interesting to compare the dynamical stability of observed
  and predicted food webs, as well as the topological stability that we
  have focused on.
\end{enumerate}
\end{abstract}

@@

\hypertarget{background-on-anthropogenic-changes-and-its-impact-on-food-web}{%
\subsection{Background on anthropogenic changes and its impact on food
web}\label{background-on-anthropogenic-changes-and-its-impact-on-food-web}}

Anthropogenic changes such as climate change and habitat destruction are
a threat to biodiversity, and can lead to food web collapse (Ullah et
al. 2018). This collapse in the food web is due to cascades of secondary
extinctions in a food web because of primary loss of species, for
example due to habitat destruction. The rate of collapse of these
predicted food webs is dependent on the structure and complexity of
those food webs (Jennifer A. Dunne, Williams, and Martinez 2002b;
Jennifer A. Dunne and Williams 2009). Therefore, research focused on
cascading secondary extinctions also known as `community viability
analysis' have been performed extensively in the past few decades
(Jennifer A. Dunne, Williams, and Martinez 2002b; Jennifer A. Dunne and
Williams 2009; Berg et al. 2011; Ebenman, Law, and Borrvall 2004;
Ebenman and Jonsson 2005).

\hypertarget{briefly-explain-work-done-by-jennifer-dunne-on-food-web-op-robustness-to-primary-extinctions-and-also-what-has-been-done-since-particularly-concerning-the-importance-of-connectance}{%
\subsection{Briefly explain work done by Jennifer Dunne on food web
{[}OP{]} robustness to primary extinctions, and also what has been done
since, particularly concerning the importance of
connectance}\label{briefly-explain-work-done-by-jennifer-dunne-on-food-web-op-robustness-to-primary-extinctions-and-also-what-has-been-done-since-particularly-concerning-the-importance-of-connectance}}

Simulation of primary species loss has been conducted in observed food
webs and model food webs from terrestrial and aquatic ecosystems where
robustness was measured in terms of secondary extinctions (Jennifer A.
Dunne, Williams, and Martinez 2002b; Jennifer A. Dunne and Williams
2009). These studies showed that the robustness \textbf{of the food
webs} increases with food web connectance. Also, the removal of the most
connected species cause considerably more secondary extinctions than
random removals of species (Jennifer A. Dunne, Williams, and Martinez
2002a; Sol'e and Montoya 2001). These studies provide an alternate
solution to investigate the impact of primary extinctions in a food web
when canonical experiments in natural ecosystems are not possible.

Along with robustness based on topological structure of a food web,
robustness based on the food web dynamics has been studied as well
(reference). Topological approaches only require the food web structure
whereas dynamical approaches also require the temporal dynamics of the
food web along with the food web structure. For example: Williams (2008)
combined models of network structure with models of bionergetic dynamics
to study the role of food web topology and nonlinear dynamics on species
coexistence in complex ecological networks.

{[}OP{]} Explain (either here or in the methods) difference between
topological and abundance (given by dynamics) based criteria for a
secondary extinction occurring.

\hypertarget{op-state-what-is-the-problem}{%
\subsection{{[}OP{]} State what is the
problem\ldots{}}\label{op-state-what-is-the-problem}}

A key assumption of the observed food webs is that the food webs are
very well sampled i.e.~all the links that in reality can occur are
represented. However, it is known that not all food webs are very well
sampled and then do not represent all of the feeding links that occur
(Caron et al. 2022; Patonai and Jord'an 2017; Jordano 2016). Some rare
trophic links require more sampling effort as compared to others whereas
some trophic links remain unobserved because of linkage constraints
irrespective of sufficient sampling effort (Jordano 2016). One solution
to this problem is to use a food web model such as Allometric Diet
Breadth Model (ADBM) to predict which are the missing links, and to then
measure the robustness of the predicted food web. ADBM is a mechanistic
model constructed using rules based on body sizes of prey and predator
where trophic interactions satisfying those rules would be predicted by
the model which at the same time are perhaps not observed because those
interactions are rare. However, this solution is not infallible, as it
is likely that the food web model might still miss some links, and also
may predict some links that could not, in fact occur.

\hypertarget{what-we-do-in-this-study}{%
\subsection{What we do in this study}\label{what-we-do-in-this-study}}

In our study, we investigate the robustness of the ADBM predicted food
webs as compared to the observed food webs and quantify the effect of
overestimation of connectance on the robustness of these predicted food
webs. We do this by simulating primary species loss in 12 food webs
predicted from the ADBM to quantify the secondary loss of extinctions.
We use three different approaches of species removal: (i) most connected
species, (ii) random species and (iii) least connected species to
understand if the outcome varies among these approaches.

It is crucial to investigate the implication of this consistent
overestimation of connectance in the robustness of predicted food webs.
We expect that the ADBM predicted food webs would be more robust as
compared to the observed food webs, and for the greater robustness to be
related to the amount by which the ADBM overestimates connectance. In
this study, we simulate primary species loss in 12 food webs predicted
from the ADBM to quantify the secondary loss of extinctions. We use
three different approaches of species removal: (i) most connected
species, (ii) random species and (iii) least connected species to
understand if the outcome varies among these approaches.

@@

{[}OP{]} \#\# Provide overview of the methods

In the upcoming sections, we present a detailed account of the
implementation of simulation of primary extinctions for three different
scenarios on 12 food webs predicted by the ADBM from wide variety of
ecosystems, and compute the resultant secondary extinctions. We then
compute a robustness metric to quantify the robustness of those
predicted food webs.

@@

The allometric diet breadth model (ADBM) is based on optimal foraging
theory, specifically the contingency model (MacArthur and Pianka 1966).
The ADBM predicts the set of prey species a consumer should feed upon to
maximise its rate of energy intake (Petchey et al. 2008). The foraging
variables in the model are: energy content of prey, handling times of
the predator on prey, space clearance rate of predator on prey, and prey
densities. All are derived from the body sizes of the species via
allometries.

@@

The observed food webs that we fit the ADBM to belong to marine,
freshwater and terrestrial ecosystems (Table \ref{fig:tab_1}). The
observed connectance of these food webs is from 0.03 to 0.24 and there
are 29 to 239 species. The food webs contain primary producers,
herbivores, carnivores, parasites, and parasitoids. They also contain
various types of feeding interactions, including predation, herbivory,
bacterivory, parasitism, pathogenic, and parasitoid.

@@

\caption{\label{tab:unnamed-chunk-1}\label{fig:tab_1}Information about the food webs predicted using the ADBM.}
\centering
\resizebox{\linewidth}{!}{
\fontsize{7}{9}\selectfont
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}|>{\raggedright\arraybackslash}p{8em}|l|l|l|l|>{}p{8em}|>{}p{8em}|>{}p{8em}}
\hline
Common food web name (Original Publication) & Predation matrix source & General ecosystem & Number of species & Observed connectance & 95\% prediction interval of predicted connectance  (Gupta et al. (2022))\\
\hline
Benguela Pelagic (Yodzis 1998) & Brose et al. (2005) & Marine & 30 & 0.21 & 0.26 - 0.59\\
\hline
Broadstone Stream (taxonomic aggregation) (Woodward and Hildrew 2001; Woodward
et al. 2005) & Brose et al. (2005) & Freshwater & 29 & 0.19 & 0.18 - 0.72\\
\hline
Broom (Memmott et al. 2000) & Brose et al. (2005) & Terrestrial & 60 & 0.03 & 0.12 - 0.89\\
\hline
Capinteria (Lafferty et al. 2006) & Hechinger et al. (2011) & Marine (Salt Marsh) & 88 & 0.08 & 0.11 - 0.80\\
\hline
Caricaie Lakes (Cattin et al. 2004) & Brose et al. (2005) & Freshwater & 158 & 0.05 & 0.11 - 0.81\\
\hline
Grasslands (Dawah et al. 1995) & Brose et al. (2005) & Terrestrial & 65 & 0.03 & 0.03 - 0.44\\
\hline
Mill Stream (Ledger, Edwards, Woodward unpublished) & Brose et al. (2005) & Freshwater & 80 & 0.06 & 0.08 - 0.60\\
\hline
Skipwith Pond (Warren 1989) & Brose et al. (2005) & Freshwater & 71 & 0.07 & 0.17 - 0.90\\
\hline
Small Reef (Opitz 1996 Table 8.6.2) & Alyssa R. Cirtwill and Anna Eklöf (2018) & Marine (Reef) & 239 & 0.06 & 0.07 - 0.66\\
\hline
Tuesday Lake (Jonsson et al. 2005) & Brose et al. (2005) & Freshwater & 73 & 0.08 & 0.09 - 0.57\\
\hline
Ythan (Emmerson and Raffaelli 2004) & Alyssa R. Cirtwill and Anna Eklöf (2018) & Marine (Estuarine) & 85 & 0.04 & 0.13 - 0.84\\
\hline
Broadstone Stream (size aggregation) (Woodward
et al. 2010) & Guy Woodward. (2021) & Freshwater & 29 & 0.24 & 0.25 - 0.47\\
\hline
\end{tabular}}
\end{table}
\end{landscape}
\restoregeometry

\hypertarget{op-need-a-section-on-how-the-food-web-model-was-fit-to-the-data.-also-should-state-how-we-deal-with-use-the-uncertainty-contained-in-the-posterior-joint-distribution.}{%
\subsection{**OP: need a section on how the food web model was fit to
the data. Also should state how we deal with / use the uncertainty
contained in the posterior joint
distribution.}\label{op-need-a-section-on-how-the-food-web-model-was-fit-to-the-data.-also-should-state-how-we-deal-with-use-the-uncertainty-contained-in-the-posterior-joint-distribution.}}

\hypertarget{primary-and-secondary-extinctions}{%
\subsection{Primary and secondary
extinctions}\label{primary-and-secondary-extinctions}}

We implemented the primary species removal method from Jennifer A. Dunne
and Williams (2009) by sequentially removing species using one of three
criteria: removal of (i) the most-connected species, (ii) the
least-connected species and (iii) randomly chosen species. The
most-connected and least-connected criteria are based on the degree
(i.e.~total number of links to resources and from consumers) of species.
Given a primary removal if any remaining species lost all of their
resource species, or any cannibalistic species lost all of their
resource species except the cannibalistic links, they are removed from
the web and a secondary extinction was recorded. Secondary extinctions
may cause further secondary extinctions, which were also checked for and
recorded. One no more secondary extinctions occured, then another
primary extinctions was made, of the next appropriate species. This
process was carried out until all species were extinct from the web.

\hypertarget{calculating-robustness}{%
\subsection{Calculating robustness}\label{calculating-robustness}}

Robustness (R) of food web was quantified as the proportion of species
subjected to primary removals that resulted in a loss (i.e.~primary
removals plus secondary extinctions) of some specified proportion of the
species. In our study, we use \(R_{50}\), the number of primary
extinctions divided by the total number of species, that result in at
least 50 per cent of total species loss (Jennifer A. Dunne, Williams,
and Martinez 2002b; J. Dunne, Williams, and Martinez 2004; Jennifer A.
Dunne and Williams 2009). Therefore, if primary extinctions never cause
any secondary extinctions, the food web is maximally robust and
(\(R_{50} = 0.50\)). Whereas in a minimally robust community
(\(R_{50} = 1/S\)), since the first primary extinction causes as cascade
of secondary extinctions of at least nearly half of the species in the
food web (i.e.~at least \(S/2 - 1\)).

@@

\hypertarget{show-and-describe-the-secondary-extinction-curves}{%
\subsection{Show and describe the secondary extinction
curves}\label{show-and-describe-the-secondary-extinction-curves}}

In Fig. \ref{fig:fig_r1}, \ref{fig:fig_r2} and \ref{fig:fig_r3}, we show
the secondary extinction curves of ADBM predicted food webs and observed
food webs for 12 different food webs under three different extinction
scenarios. We found that the cumulative secondary extinction was higher
for the ADBM predicted food webs as compared to the observed food webs
for nine, nine and seven food webs.

In general, irrespective of the extinction scenarios, we found that the
cumulative secondary extinction was higher for the ADBM predicted food
webs as compared to the observed food webs for most of the food webs
\textbf{OP: this is not so clear to me. State how many cases? It seems
there are many cases where secondary extinctions are more numerous in
the observed food web: blue line above the red.}.

In the most connected extinction scenario, the cumulative secondary
extinction curve for the observed food webs rose quickly as compared to
the ADBM predicted food webs, and then reach saturation after a certain
number of primary removal of species. In some of the food webs (Fig.
\ref{fig:fig_r1} (f, g, h, i, j, k)), there were intersection between
the cumulative secondary extinction curves of ADBM predicted food webs
and that of the observed food webs. In case of the Broadstone Stream
(taxonomic aggregation) food web and the Tuesday Lake food web (Fig.
\ref{fig:fig_r1} (b and k)), the secondary extinction curves for the
ADBM food webs were higher than the observed food webs, whereas in case
of the Skipwith Pond food web (Fig. \ref{fig:fig_r1} (i)), there were no
secondary extinctions for any given number of primary removal of
species.

@@

{\centering \includegraphics[width=400px]{../results/plot_mostconnected}

@@

\caption{\label{fig:fig_r1} Cumulative secondary extinctions of species resulting from the primary **removals of the most connected species** for 12 food webs. S denotes the number of species in a food web. The cumulative secondary extinctions of species and the number of species removed have been normalised by the number of species.}\label{fig:unnamed-chunk-2}
\end{figure}

Except Broadstone Stream (taxonomic aggregation), Broadstone Stream
(size aggregation) and Tuesday Lake food webs (Fig. \ref{fig:fig_r2} (b,
c and k)), the \textbf{mean cumulative secondary extinction curves}
\textbf{OP: must be explained somewhere how/why this is a mean} for all
the other food webs predicted by the ADBM were always lower than that of
the observed food webs in the random extinction scenario. The shape of
the cumulative secondary extinctions curves varied across the food webs.

@@

{\centering \includegraphics[width=400px]{../results/plot_random}

@@

\caption{\label{fig:fig_r2} Mean cumulative secondary extinctions of species resulting from the primary **removals of random species** for 12 food webs. S denotes the number of species in a food web. The cumulative secondary extinctions of species and the number of species removed have been normalised by the number of species.}\label{fig:unnamed-chunk-3}
\end{figure}

Compared to the most connected and random extinction scenarios, the
cumulative extinction curves in the least connected extinction scenario
had very low values and were flat for most of the food webs (Fig.
\ref{fig:fig_r3}) compared to the most connected extinction and random
extinction scenarios. In most of the food webs, there was a lot of
overlap between the extinction curves of the ADBM predicted food webs
and the observed food webs.

@@

{\centering \includegraphics[width=400px]{../results/plot_leastconnected}

@@

\caption{\label{fig:fig_r3} Cumulative secondary extinctions of species resulting from the primary **removals of the least connected species** for 12 food webs. S denotes the number of species in a food web. The cumulative secondary extinctions of species and the number of species removed have been normalised by the number of species.}\label{fig:unnamed-chunk-4}
\end{figure}

\textbf{OP: write text describing the effect sizes, e.g.~how big is the
difference between observed and ADBM webs.}

\hypertarget{show-and-describe-the-following}{%
\subsection{Show and describe the
following:}\label{show-and-describe-the-following}}

>>>>

\title{Missing links and the topological robustness of food webs}

@@

\usepackage{lineno}
\usepackage {amsmath}
\setlength\parindent{24pt}
\usepackage{setspace}\doublespacing
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\begin{document}
\maketitle

\begin{abstract}
\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\tightlist
\item
  Undersampling can lead to missing trophic interactions in recorded
  food webs, with potential consequences for the perceived functioning
  and stability of the food webs. Undersampling can be compensated for
  by using food web models such as the allometric diet breadth model
  (ADBM) to predict missing links. Simultaneously, models might predict
  links which cannot occur, i.e., false positives.
\item
  Previous research shows that (i) food web robustness (the inverse of
  the number of secondary extinctions occuring due to primary
  extinctions) increases with connectance (the number of realised
  trophic links divided by the number of possible links), and (ii) that
  predicted food webs usually have greater connectance than observed
  ones. Thus we expect that predicted food webs are more robust than
  observed ones. This expectation has never, to our knowledge, been
  tested, nor has the effect size been quantified.
\item
  We fill this research gap by comparing the robustness of observed food
  webs to the robustness of food webs predicted by a model (the ADBM)
  that can account for missing links, though can also make false
  positives. We did this for 12 different food webs from a wide variety
  of ecosystems.
\item
  We found, as expected, that the predicted food webs were more robust
  than the observed food webs, and this can be attributed to the higher
  connectance of the predicted food webs. On average, for every one unit
  of increase in connectance, we found the food webs to be robust by
  0.52 units and 0.04 units for the most connected and the random
  species extinction scenarios respectively while no effect in the least
  connected species extinction scenario.
\item
  These results show that undersampling can lead to large underestimates
  of food web robustness that can be compensated for by filling in
  missing links with food web models. Nevertheless, increased
  connectance may contribute to lower dynamical stability, and so it
  would be interesting to compare the dynamical stability of observed
  and predicted food webs, as well as the topological stability that we
  have focused on.
\end{enumerate}
\end{abstract}

@@

Anthropogenic changes such as climate change and habitat destruction are
a threat to biodiversity, and can lead to food web collapse (Ullah et
al. 2018). This food web collapse is due to the cascades of secondary
extinctions in a food web because of the primary loss of species, for
example due to habitat destruction and climate change (Pimm et al. 2006;
J. A. Thomas et al. 2004; C. D. Thomas et al. 2004). Therefore, research
focused on cascading secondary extinctions also known as `community
viability analysis' have been performed extensively in the past few
decades to quantify how robust are food webs to species extinction
(Jennifer A. Dunne, Williams, and Martinez 2002b; Jennifer A. Dunne and
Williams 2009; Berg et al. 2011; Ebenman, Law, and Borrvall 2004;
Ebenman and Jonsson 2005). And it has been shown that the rate of
collapse of a food web is dependent on its structure and complexity
(Jennifer A. Dunne, Williams, and Martinez 2002b; Jennifer A. Dunne and
Williams 2009).

Simulation of primary species loss has been conducted in observed food
webs and model food webs from terrestrial and aquatic ecosystems where
robustness was measured in terms of secondary extinctions (Jennifer A.
Dunne, Williams, and Martinez 2002b; Jennifer A. Dunne and Williams
2009). These studies have shown that the robustness of the food webs
increases with food web connectance. Also, the removal of the most
connected species cause considerably more secondary extinctions than the
random removals of species (Jennifer A. Dunne, Williams, and Martinez
2002a; Sol'e and Montoya 2001). Simulation studies like these which
investigate the impact of primary extinctions in a food web to quantify
robustness based on its topological structure provide an alternate
solution to canonical experiments in natural ecosystems which are not
possible or very difficult to conduct (Jennifer A. Dunne and Williams
2009).

Along with quantifying food web robustness based on its topological
structure, studies such as Williams (2008), Brose, Williams, and
Martinez (2006) and Martinez, Williams, and Dunne (2006) have quantified
robustness based on the abundance dynamics of a food web. The
topological approach of quantifying a food web robustness only requires
the food web structure whereas the dynamical approach not only requires
the food web structure but also the temporal dynamics of abundance of
species in that food web. For example: Williams (2008) combined models
of network structure with models of bioenergetic dynamics to study the
role of food web topology and nonlinear dynamics on species coexistence
in complex ecological networks.

A key assumption of the observed food webs is that they are very well
sampled i.e.~all the links that in reality can occur are represented.
However, it is known that not all food webs are very well sampled and
then do not represent all of the feeding links that occur (Caron et al.
2022; Patonai and Jord'an 2017; Jordano 2016). Some rare trophic links
require more sampling effort as compared to others whereas some trophic
links remain unobserved because of linkage constraints irrespective of
sufficient sampling effort (Jordano 2016). Previous studies such as
Caron et al. (2022) and Gupta, Furrer, and Petchey (2022) have shown
that the predicted food webs from these models usually have greater
connectance than the observed ones. Therefore, one solution to
compensate for undersampling is to use a food web model such as the
Allometric Diet Breadth Model (ADBM) to predict the missing links, and
to then measure the robustness of the predicted food web. ADBM is a
mechanistic model constructed using foraging rules based on the body
sizes of prey and predator where trophic interactions satisfying those
rules would be predicted by the model which are perhaps not observed
because those interactions are rare. However, this solution is not
infallible, as it is likely that the food web model might still miss
some links, and also may predict some links that could not, in fact
occur.

In our study, we investigate the topological robustness of the ADBM
predicted food webs and compare it to that of the observed food webs,
and quantify the effect of undersampling i.e overestimation of
connectance on the robustness of these predicted food webs. We expect
that the ADBM predicted food webs to be more robust as compared to the
observed food webs, and for the greater robustness to be related to the
amount by which the ADBM overestimates connectance. We do this by
simulating primary species loss in 12 food webs predicted from the ADBM
to quantify the secondary loss of extinctions. We use three different
approaches of species removal: (i) most connected species, (ii) random
species and (iii) least connected species to understand if the outcome
varies among these approaches.

@@

In the upcoming sections, we present a detailed account of the
implementation of simulation of primary extinctions for three different
extinction scenarios on 12 food webs predicted by the ADBM from a wide
variety of ecosystems and compute the resultant secondary extinctions.
We then compute a robustness metric to quantify the robustness of those
predicted food webs and compare them against that of the observed food
webs.

@@

The allometric diet breadth model (ADBM) is based on optimal foraging
theory, specifically the contingency model (MacArthur and Pianka 1966).
We chose this model because the ADBM can likely predict missing links in
the predicted food webs because it consistently overestimated
connectance in the predicted food webs as shown by our study in Gupta,
Furrer, and Petchey (2022). The ADBM predicts the set of prey species a
consumer should feed upon to maximise its rate of energy intake (Petchey
et al. 2008). The foraging variables used in the model are the energy
content of prey, handling times of the predator on prey, space clearance
rate i.e.~how fast a predator searches space, and prey densities. All of
these variables are derived from the allometric scaling relationship
using the body sizes of species. More details on the foraging rules
defined in the ADBM and ADBM's predictive power across different food
webs can be found in Petchey et al. (2008).

@@

The observed food webs that we fit the ADBM to belong to marine,
freshwater and terrestrial ecosystems (Table \ref{fig:tab_1}). These
food webs contain primary producers, herbivores, carnivores, parasites,
and parasitoids and also contain various types of feeding interactions,
including predation, herbivory, bacterivory, parasitism and pathogenic.
The observed connectance of these food webs varies from 0.03 to 0.24 and
the number of species varies from 29 to 239 species.

The goodness of fit of the ADBM's predictions depends on the interaction
types in the food webs with predictions that are more size-structured
such as aquatic and herbivory interactions being better predicted when
compared to less size-structured ones such as parasitoid and terrestrial
herbivory ones (Petchey et al. 2008).

@@

\caption{\label{tab:unnamed-chunk-1}\label{fig:tab_1}Information about the food webs predicted using the ADBM.}
\centering
\resizebox{\linewidth}{!}{
\fontsize{7}{9}\selectfont
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}|>{\raggedright\arraybackslash}p{8em}|l|l|l|l|>{}p{8em}|>{}p{8em}|>{}p{8em}}
\hline
Common food web name (Original Publication) & Predation matrix source & General ecosystem & Number of species & Observed connectance & 95\% prediction interval of predicted connectance  (Gupta et al. 2022)\\
\hline
Benguela Pelagic (Yodzis 1998) & Brose et al. (2005) & Marine & 30 & 0.21 & 0.26 - 0.59\\
\hline
Broadstone Stream (taxonomic aggregation) (Woodward and Hildrew 2001; Woodward
et al. 2005) & Brose et al. (2005) & Freshwater & 29 & 0.19 & 0.18 - 0.72\\
\hline
Broom (Memmott et al. 2000) & Brose et al. (2005) & Terrestrial & 60 & 0.03 & 0.12 - 0.89\\
\hline
Capinteria (Lafferty et al. 2006) & Hechinger et al. (2011) & Marine (Salt Marsh) & 88 & 0.08 & 0.11 - 0.80\\
\hline
Caricaie Lakes (Cattin et al. 2004) & Brose et al. (2005) & Freshwater & 158 & 0.05 & 0.11 - 0.81\\
\hline
Grasslands (Dawah et al. 1995) & Brose et al. (2005) & Terrestrial & 65 & 0.03 & 0.03 - 0.44\\
\hline
Mill Stream (Ledger, Edwards, Woodward unpublished) & Brose et al. (2005) & Freshwater & 80 & 0.06 & 0.08 - 0.60\\
\hline
Skipwith Pond (Warren 1989) & Brose et al. (2005) & Freshwater & 71 & 0.07 & 0.17 - 0.90\\
\hline
Small Reef (Opitz 1996 Table 8.6.2) & Alyssa R. Cirtwill and Anna Eklöf (2018) & Marine (Reef) & 239 & 0.06 & 0.07 - 0.66\\
\hline
Tuesday Lake (Jonsson et al. 2005) & Brose et al. (2005) & Freshwater & 73 & 0.08 & 0.09 - 0.57\\
\hline
Ythan (Emmerson and Raffaelli 2004) & Alyssa R. Cirtwill and Anna Eklöf (2018) & Marine (Estuarine) & 85 & 0.04 & 0.13 - 0.84\\
\hline
Broadstone Stream (size aggregation) (Woodward
et al. 2010) & Guy Woodward. (2021) & Freshwater & 29 & 0.24 & 0.25 - 0.47\\
\hline
\end{tabular}}
\end{table}
\end{landscape}
\restoregeometry

\hypertarget{model-parameterisation-using-approximate-bayesian-computation}{%
\subsection{Model parameterisation using approximate Bayesian
computation}\label{model-parameterisation-using-approximate-bayesian-computation}}

The ADBM was parameterised using approximate Bayesian computation (ABC)
where a set of parameter values were sampled from the prior
distributions. That set of parameter values was either accepted or
rejected based on how close the predicted food web is to the observed
food web using an accuracy metric -- true skill statistic (TSS). The
accepted parameter values then formed a posterior distribution. Further,
prediction intervals of the true skill statistic and connectance of the
predicted food webs were computed. In our study, we considered food webs
where the predicted connectance lay within the 95\% prediction interval.
A detailed explanation of the parameterisation method can be found in
Gupta, Furrer, and Petchey (2022).

\hypertarget{extinction-scenarios-and-robustness}{%
\subsection{Extinction scenarios and
robustness}\label{extinction-scenarios-and-robustness}}

We implemented the primary species removal method from Jennifer A. Dunne
and Williams (2009) by sequentially removing species using one of the
three criteria: removal of (i) the most-connected species, (ii) the
least-connected species and (iii) randomly chosen species. The
most-connected and least-connected criteria are based on the degree
(i.e.~the total number of links to resources and from consumers) of
species. We considered these mentioned criteria because the random
extinction scenario takes into account all the theoretically possible
extinction sequences of species that can occur in a food web and the
extinction of most-connected species and least-connected species takes
into account the two opposite extreme scenarios. These extinction
scenarios have been widely used in studying species extinctions and
collapse of food webs and other networks (Jennifer A. Dunne, Williams,
and Martinez 2002b; Sol'e and Montoya 2001; J. Dunne, Williams, and
Martinez 2004; Jennifer A. Dunne and Williams 2009; Albert and Barab'asi
2002).

Given a primary removal of species in a food web if any remaining
species lost all of their resource species, or any cannibalistic species
lost all of their resource species except the cannibalistic links, they
are removed from the web and a secondary extinction was recorded.
Secondary extinctions may cause further secondary extinctions, which
were also checked for and recorded. Once no more secondary extinctions
occurred, then another primary extinction was made, of the next
appropriate species depending on the extinction scenario. This process
was carried out until all the species were extinct from the food web.

The robustness (R) of a food web was defined as the proportion of
species subjected to primary removals that resulted in a set of
extinction (i.e.~primary removals plus secondary extinctions) of some
specified proportion of the species. In our study, we use \(R_{50}\),
the number of primary extinctions divided by the total number of
species, which results in at least 50\% of total species loss (Jennifer
A. Dunne, Williams, and Martinez 2002b; J. Dunne, Williams, and Martinez
2004; Jonsson et al. 2015; Jennifer A. Dunne and Williams 2009).
Therefore, if primary extinctions never cause any secondary extinctions,
the food web is maximally robust and (\(R_{50} = 0.50\)). Whereas in a
minimally robust community (\(R_{50} = 1/S\)), the first primary
extinction causes a cascade of secondary extinctions of at least nearly
half of the species in the food web (i.e.~at least \(S/2 - 1\)).

\hypertarget{simulating-species-extinctions}{%
\subsection{Simulating species
extinctions}\label{simulating-species-extinctions}}

First, we simulated primary species loss in food webs predicted by the
ADBM which had the maximum true skill statistics and compared it to
primary species loss in observed food webs. Second, to take into account
the uncertainty in robustness in the ADBM predicted food webs we
simulated primary species loss and thereby computed robustness for all
the ADBM predicted food webs corresponding to the 95\% prediction
interval of the predicted connectance. In the case of the random
extinction scenario, we simulated 1000 random extinction sequences in a
single ADBM predicted food web.

\hypertarget{statistical-analysis}{%
\subsection{Statistical analysis}\label{statistical-analysis}}

In the random extinction scenario, we computed robustness \(R_{50}\) for
all 1000 independent random extinction sequences and calculated the
median as a summary statistics to quantify the average robustness of a
single food web. To quantify the effect of undersampling
i.e.~overestimation of connectance we compute the ratio of the
difference in normalised robustness between the ADBM predicted food webs
and observed food webs to the difference in their normalised
connectance, where normalisation was performed by dividing the variables
by their maximum possible values (I.e 0.5 for \(R_{50}\) and 1 for
connectance). However, we did not perform any statistical significance
test because we work with simulated food webs and therefore the p-values
of these tests are influenced by the number of model simulations (White
et al. 2014).

@@

We first present the secondary extinction curves of the ADBM predicted
food webs which had the maximum true skill statistics and the observed
food webs for 12 food webs under three different extinction scenarios.
We then compare the robustness of all the ADBM predicted food webs
within the 95\% prediction interval against that of the observed food
webs to take into account uncertainty in the robustness across food webs
predictions. Finally we quantify the effect of overestimation of
connectance on the difference in their robustness estimates.

\hypertarget{secondary-extinctions}{%
\subsection{Secondary extinctions}\label{secondary-extinctions}}

In the most-connected extinction scenario, the cumulative secondary
extinction curves started to rise steeply at a lower number of primary
species removal in the observed food webs as compared to the ADBM
predicted food webs for nine food webs (Fig. \ref{fig:fig_r1} (a, c, d,
e, f, g, h, j, l)). However there were higher number of cumulative
secondary extinctions occurring in the ADBM predicted food webs when
compared to that of the observed food webs at a high number of primary
species removal (Fig. \ref{fig:fig_r1} (a, f, g, h, k)). In the Skipwith
Pond food web, there were no secondary extinctions for any number of
primary removal of species (Fig. \ref{fig:fig_r1} (i)), whereas in the
Broadstone Stream (taxonomic aggregation) food web the same was true
only for the observed food web but in the ADBM predicted food web there
was a steep rise in the cumulative secondary extinctions (Fig.
\ref{fig:fig_r1} (b)).

@@

{\centering \includegraphics[width=400px]{../results/plot_mostconnected_maxTSS}

@@

\caption{\label{fig:fig_r1} Cumulative secondary extinctions of species resulting from the primary \textbf{removals of the most connected species} in the ADBM predicted food webs corresponding to the maximum TSS and observed food webs. S denotes the number of species in a food web. The cumulative secondary extinctions of species and the number of species removed have been normalised by the number of species.}\label{fig:unnamed-chunk-2}
\end{figure}

In Fig. \ref{fig:fig_r2}, we present the cumulative secondary
extinctions in the ADBM predicted and the observed food webs for five
(out of 1000) independent random extinction sequences to show
qualitative differences between their secondary extinction curves. The
secondary extinctions in the ADBM predicted food webs were more abrupt
than that in the observed food webs.

@@

{\centering \includegraphics[width=400px]{../results/plot_ra_extlines_maxTSS}

@@

\caption{\label{fig:fig_r2} Cumulative secondary extinctions of species resulting from the primary \textbf{removals of random species} in the ADBM predicted food webs corresponding to the maximum TSS and observed food webs for five independent random extinction sequences. S denotes the number of species in a food web. The cumulative secondary extinctions of species and the number of species removed have been normalised by the number of species.}\label{fig:unnamed-chunk-3}
\end{figure}

Compared to the most-connected and random extinction scenarios, there
were fewer secondary extinctions in the least-connected extinction
scenario and therefore the secondary extinction curves were flat for
most of the food webs (Fig. \ref{fig:fig_r3}). In some of the food webs,
the extinction curves of the ADBM predicted food webs overlapped with
the observed food webs (Fig. \ref{fig:fig_r3} (b, c, g, h, i, k, l)\}).
Contrary to the most-connected scenario, there was a very high number of
secondary extinctions occurring at very low number of primary species
removal (Fig. \ref{fig:fig_r3} (a, d, e, f, j)).

@@

{\centering \includegraphics[width=400px]{../results/plot_leastconnected_maxTSS}

@@

\caption{\label{fig:fig_r3} Cumulative secondary extinctions of species resulting from the primary \textbf{removals of the least connected species} in the ADBM predicted food webs corresponding to the maximum TSS and observed food webs. S denotes the number of species in a food web. The cumulative secondary extinctions of species and the number of species removed have been normalised by the number of species.}\label{fig:unnamed-chunk-4}
\end{figure}

\hypertarget{robustness}{%
\subsection{Robustness}\label{robustness}}
